Ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2018-01-24 manuscript/usses_water_ms.tex
adds PDF version of figs for resubmission

We conducted our precipitation manipulation experiment in a sagebrush
steppe community at the USDA-ARS Sheep Experiment Station (USSES) near
Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above
sea level. The plant community is dominated by the shrub
\emph{Artemisia tripartita},
\new{the perennial forb \emph{Balsamorhiza sagittata}}, and three
perennial bunchgrasses, \emph{Pseudoroegneria spicata},
\emph{Poa secunda}, and \emph{Hesperostipa comata}
\new{(see Appendix 1 for rank abundance curves)}. During the period of
our experiment (2011 -- 2016), mean annual precipitation was 250 mm
year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
and 1932, range scientists at the USSES established 26 permanent 1
m\(^2\) quadrats to track vegetation change over time. In 2007, we
relocated 14 of the original quadrats in permenanent livestock
exclosures, \new{which we used} as control plots (i.e.~ambient
precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because collecting demographic data is time consuming and it was already being collected in these plots for other studies.}

@@

\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start unitl April) and for a string of days in 2013 during which the soil moisture probes failed to take readings.
To fill in these gaps, we used a version of the SOILWAT soil moisture model}
\citep{Sala1992a}
\new{that has been specifically designed for semiarid shrublands and grasslands}
\citep{Bradford2014}.
\new{The model was parameterized using generic sagebrush steppe parameters and local soil texture, soil bulk density, and precipitation data}
\citep{Kleinhesselink2017b}.
\new{We used the locally parameterized SOILWAT model to generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig. 2B and Appendix 2).
SOILWAT predicted daily soil moisture under ambient conditions similar to our control plots.
We applied the same statistical model and procedure described above to estimate soil moisture in drought and irrigation plots based on control plot conditions.}

@@

\begin{table}[ht]
\centering
\caption{Summary statistics from the posterior distributions of coefficients for each treatment ($\beta$ coefficients in equation 1). The `Intercept' and `Slope' summaries reported here for drought and irrigation are from the posterior distributions of the intercept and slope for the control treatment plus the offsets for each treatment. Posterior distributions of the offsets are in Figure 3A.} 
\begingroup\normalsize
\begin{tabular}{llrrrr}
  \hline
Coefficient & Treatment & Posterior Mean & Posterior Median & Lower 95\% BCI & Upper 95\% BCI \\ 
  \hline
Intercept & Control & 0.14 & 0.14 & -1.24 & 1.50 \\ 
  Intercept & Drought & 0.67 & 0.65 & -2.01 & 3.36 \\ 
  Intercept & Irrigation & -0.39 & -0.34 & -3.51 & 2.48 \\ 
  Slope & Control & 0.53 & 0.52 & -1.44 & 2.66 \\ 
  Slope & Drought & 1.02 & 1.00 & -1.04 & 3.25 \\ 
  Slope & Irrigation & 0.56 & 0.55 & -1.43 & 2.71 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}\begin{table}[ht]
\centering
\caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis' tests whether treatments form unique clusters in multidimensional space; `betadisper' tests whether treatments have similar dispersion. For both tests, \emph{P} values greater than 0.05 indicate there is no support for the hypothesis that the treatments differ.} 
\begingroup\normalsize
\begin{tabular}{rlrrrr}
  \hline
Year & Test & n & d.f. & \emph{F} & \emph{P} \\ 
  \hline
2011 & adonis &  22 &   2 & 1.04 & 0.43 \\ 
  2011 & betadisper &  22 &   2 & 2.22 & 0.14 \\ 
  2012 & adonis &  22 &   2 & 1.09 & 0.32 \\ 
  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
  2013 & adonis &  22 &   2 & 1.26 & 0.12 \\ 
  2013 & betadisper &  22 &   2 & 0.43 & 0.66 \\ 
  2014 & adonis &  22 &   2 & 0.96 & 0.58 \\ 
  2014 & betadisper &  22 &   2 & 0.36 & 0.70 \\ 
  2015 & adonis &  22 &   2 & 1.08 & 0.33 \\ 
  2015 & betadisper &  22 &   2 & 1.96 & 0.17 \\ 
  2016 & adonis &  22 &   2 & 1.12 & 0.26 \\ 
  2016 & betadisper &  22 &   2 & 0.27 & 0.76 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}

>>>>

We conducted our precipitation manipulation experiment in a sagebrush
steppe community at the USDA-ARS Sheep Experiment Station (USSES) near
Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above
sea level. The plant community is dominated by the shrub
\emph{Artemisia tripartita},
\new{the perennial forb \emph{Balsamorhiza sagittata}}, and three
perennial bunchgrasses, \emph{Pseudoroegneria spicata},
\emph{Poa secunda}, and \emph{Hesperostipa comata}
\new{(see Appendix 1 for rank abundance curves)}. During the period of
our experiment (2011 -- 2016), mean annual precipitation was 250 mm
year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
and 1932, range scientists at the USSES established 26 permanent 1
m\(^2\) quadrats to track vegetation change over time. In 2007, we
located 14 of the original quadrats in permenanent livestock exclosures,
\new{which we used} as control plots (i.e.~ambient precipitation) in the
experiment described below.
\new{We used the original plots as our controls because collecting demographic data is time consuming and it was already being collected in these plots for other studies.}

@@

\new{Following the above procedure, we still lacked soil moisture data for March 2012 (observations did not start until April 2012) and for a string of days in 2013 during which the soil moisture probes failed to take readings.
To fill in these gaps, we used a version of the SOILWAT soil moisture model}
\citep{Sala1992a}
\new{that has been specifically designed for semiarid shrublands and grasslands}
\citep{Bradford2014}.
\new{The model was parameterized using generic sagebrush steppe parameters and local soil texture, soil bulk density, and precipitation data}
\citep{Kleinhesselink2017b}.
\new{We used the locally parameterized SOILWAT model to generate daily soil moisture predictions for the duration of the experiment, but only used SOILWAT predictions where there were gaps in our data (Fig. 2B and Appendix 2).
SOILWAT predicted daily soil moisture under ambient conditions similar to our control plots.
We applied the same statistical model and procedure described above to estimate soil moisture in drought and irrigation plots based on control plot conditions.}

@@

\begin{table}[ht]
\centering
\caption{Summary statistics from the posterior distributions of coefficients for each treatment ($\beta$ coefficients in equation 1). The `Intercept' and `Slope' summaries reported here for drought and irrigation are from the posterior distributions of the intercept and slope for the control treatment plus the offsets for each treatment. Posterior distributions of the offsets are in Figure 3A.} 
\begingroup\normalsize
\begin{tabular}{llrrrr}
  \hline
Coefficient & Treatment & Posterior Mean & Posterior Median & Lower 95\% BCI & Upper 95\% BCI \\ 
  \hline
Intercept & Control & 0.14 & 0.14 & -1.24 & 1.50 \\ 
  Intercept & Drought & 0.67 & 0.65 & -2.01 & 3.36 \\ 
  Intercept & Irrigation & -0.39 & -0.34 & -3.51 & 2.48 \\ 
  Slope & Control & 0.53 & 0.52 & -1.44 & 2.66 \\ 
  Slope & Drought & 1.02 & 1.00 & -1.04 & 3.25 \\ 
  Slope & Irrigation & 0.56 & 0.55 & -1.43 & 2.71 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}\begin{table}[ht]
\centering
\caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis' tests whether treatments form unique clusters in multidimensional space; `betadisper' tests whether treatments have similar dispersion. For both tests, \emph{P} values greater than 0.05 indicate there is no support for the hypothesis that the treatments differ.} 
\begingroup\normalsize
\begin{tabular}{rlrrrr}
  \hline
Year & Test & n & d.f. & \emph{F} & \emph{P} \\ 
  \hline
2011 & adonis &  22 &   2 & 1.04 & 0.41 \\ 
  2011 & betadisper &  22 &   2 & 2.22 & 0.14 \\ 
  2012 & adonis &  22 &   2 & 1.09 & 0.32 \\ 
  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
  2013 & adonis &  22 &   2 & 1.26 & 0.11 \\ 
  2013 & betadisper &  22 &   2 & 0.43 & 0.66 \\ 
  2014 & adonis &  22 &   2 & 0.96 & 0.55 \\ 
  2014 & betadisper &  22 &   2 & 0.36 & 0.70 \\ 
  2015 & adonis &  22 &   2 & 1.08 & 0.32 \\ 
  2015 & betadisper &  22 &   2 & 1.96 & 0.17 \\ 
  2016 & adonis &  22 &   2 & 1.12 & 0.26 \\ 
  2016 & betadisper &  22 &   2 & 0.27 & 0.76 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}
